\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-26 coqimpl.tex
comments on partial recursors

\textbf{Propositional Partial Recursor.}
Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
not expressive enough to really fulfill \ref{langdesign-req2}. For
example, we cannot prove the constructors are injective, i.e.,
\mintinline{Coq}{self_B.tt ≠ self_B.ff}. In the original setting, to
prove injectivity of constructors of a (vanilla) inductive type, we only
require a recursor and its computational rules. However, as we pointed
out earlier, the recursor of an inductive type cannot be exposed into
the context for the sake of extensibility. Hence, we need to alter the
formulation of the recursor: we need to have a better
characterization of ``extensible inductive types'' exported in the
compiled module type as a part of the interface for the following fields
to use.

The simplest idea is to allow partiality---i.e., returning
\mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
We call this \textit{propositional partial recursor}.
For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}
\begin{minted}[escapeinside=@@]{Coq}
b_prec : forall P : b -> Set, P tt -> P ff
    -> forall x : b, option (P x).
\end{minted}
with appropriate computational axioms. The partiality doesn't break
extensibility (i.e. all future extensions of "b" can support this
"b_prec").

>>>>

\textbf{Propositional Partial Recursor.}
Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
not expressive enough to really fulfill \ref{langdesign-req2}. For
example, we cannot prove the constructors are injective, i.e.,
\mintinline{Coq}{self_B.tt ≠ self_B.ff}.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only? My confusion after reading these paras is that I still don't know why partial recursors are needed.}
In the original setting, to
prove injectivity of constructors of a (vanilla) inductive type, we only
require a recursor and its computational rules. However, as we pointed
out earlier, the recursor of an inductive type cannot be exposed into
the context for the sake of extensibility. Hence, we need to alter the
formulation of the recursor: we need to have a better
characterization of ``extensible inductive types'' exported in the
compiled module type as a part of the interface for the following fields
to use.

The simplest idea is to allow partiality---i.e., returning
\mintinline{Coq}{option T} instead of the original \mintinline{Coq}{T}.
We call this \textit{propositional partial recursor}.\YZ{What is propositional about partial recursors? I suppose it has to do with the computational axioms?}
For example, the propositional partial recursor for \mintinline{Coq}{B.b} is\YZ{Mention that partial recursors can be automatically generated}
\begin{minted}[escapeinside=@@]{Coq}
b_prec : forall P : b -> Set, P tt -> P ff
    -> forall x : b, option (P x).@\YZ{Not clear what it takes to define/prove partial recursors. The reader needs to be told if they require breaking abstraction like recursors do.}@
\end{minted}
with appropriate computational axioms. The partiality doesn't break
extensibility (i.e., all future extensions of "b" can support this
"b_prec").
